---
title: "PA1_template.Rmd"
author: "FLat"
date: "3 11 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path='O:\\Workspace Latzelsperger\\Education and training\\Programming\\Coursera\\Reproducible Research\\Results/')

dataDir <- "O:\\Workspace Latzelsperger\\Education and training\\Programming\\Coursera\\Reproducible Research\\Data"
codeDir <- "O:\\Workspace Latzelsperger\\Education and training\\Programming\\Coursera\\Reproducible Research\\Code"
resulstDir <- "O:\\Workspace Latzelsperger\\Education and training\\Programming\\Coursera\\Reproducible Research\\Results"
```

## Reproducible reseach - Assignment 1
# Load libraries
```{r}
library(data.table)
library(stats)
library(ggplot2)
library(gridExtra)
```

# Pull data
```{r}
setwd(dataDir)
data <- as.data.table(read.csv("activity.csv"))

```

# Q1: What is mean total number of steps taken per day?
```{r}
nStepsDay <- aggregate(steps ~ date, data, FUN=sum, na.rm=TRUE)
meanStepsDay <- mean(nStepsDay$steps)
medianStepsDay <- median(nStepsDay$steps)
```

Mean
```{r}
meanStepsDay
```

Median
```{r}medianStepsDay
medianStepsDay
```

Histogram
```{r}
hist <- ggplot(nStepsDay) +
      geom_histogram(aes(x=steps), bins=5) +
      ylim(0, 40) +
      labs(title="Total steps per day", x="Total number of steps", y="Frequency")
hist
```

# Q2: What is the average daily activity pattern?
```{r}
meanStepsInterval <- aggregate(steps ~ interval, data, FUN = mean, na.rm = TRUE)

timeSeries <- ggplot(meanStepsInterval) +
      geom_line(aes(x=interval, y=steps)) +
      labs(title="Average number of steps per interval", x="Intervals", y="Average number of steps")
timeSeries
```

The following interval contains the maximum number of steps
```{r}
which(meanStepsInterval$steps==max(meanStepsInterval$steps))
meanStepsInterval[which(meanStepsInterval$steps==max(meanStepsInterval$steps)), "interval"]
```


# Q3: Imputing missing values
Total number of missing values
```{r}
sum(!complete.cases(data))
```

Strategy for imputing NAs:
First iteration: replace NAs with daily means
Second iteration: replace remaining NAs with global mean

```{r}
# Replace NAs with daily mean
partialMeanStepsDay <- aggregate(steps ~ date, data, FUN=mean, na.rm=TRUE)
names(partialMeanStepsDay)[2] <- "meanSteps"
data <- merge(data, partialMeanStepsDay, by="date", all=TRUE)
data[is.na(data$steps), steps := meanSteps]
data[is.na(data$steps), steps := mean(data$steps, na.rm=TRUE)]
data[, meanSteps := NULL]
setcolorder(data, c("steps", "date", "interval"))

nStepsDay <- aggregate(steps ~ date, data, FUN=sum, na.rm=TRUE)
meanStepsDay <- mean(nStepsDay$steps)
medianStepsDay <- median(nStepsDay$steps)
```

Mean after imputing NAs
```{r}
meanStepsDay
```

Median after imputing NAs
```{r}
medianStepsDay
```

```{r}
hist2 <- ggplot(nStepsDay) +
      geom_histogram(aes(x=steps), bins=5) +
      ylim(0, 40) +
      labs(title="Total steps per day (NAs imputed)", x="Total number of steps", y="Frequency")
hist2

grid.arrange(hist, hist2, nrow = 1)
```

Daily number of steps increases after imputing NAs.


# Q4: Are there differences in activity patterns between weekdays and weekends?
```{r}
data[, date := as.POSIXct.Date(data$date)]
data[, dayType := weekdays(data$date)]
data[dayType != "Samstag" & dayType != "Sonntag", dayType := "weekday"]
data[dayType == "Samstag" | dayType == "Sonntag", dayType := "weekend"]
data[, dayType := as.factor(data$dayType)]

meanStepsDayType <- aggregate(steps ~ interval + dayType, data = data, FUN = mean)

timeseries2 <- ggplot(meanStepsDayType, aes(interval, steps)) +
      geom_line() + 
      facet_grid(dayType~.) + 
      labs(x="Intervals", y="Average steps", title = "Average number of steps per day            during the week and on weekends")
timeseries2
```


